\chapter{Conclusion and Future Works}

\section{Conclusion}

This thesis has presented a diverse collection of research contributions spanning two broad areas: log-based anomaly detection using representation learning, and the design and analysis of image encryption and secret image sharing schemes. Together, these seven chapters address critical challenges in system security, data confidentiality, and intelligent monitoring.

Chapter~1 introduced a role-aware representation learning framework for log-based anomaly detection. By inferring latent token roles without any manual annotation or parsing, the proposed method overcomes the uniform token treatment of prior approaches such as NeuralLog. The framework achieved state-of-the-art F1-scores of 0.98--0.99 across four public datasets, demonstrating that structural and semantic information encoded through role modeling significantly enhances anomaly detection accuracy.

Chapters~2 and~3 presented chosen-plaintext cryptanalysis of two chaos-based image encryption schemes. In Chapter~2, the medical image encryption scheme of Jain et al.\ was shown to be completely breakable using only $n+1$ queries, achieving pixel-perfect reconstruction with zero maximum pixel difference. This was despite the scheme reporting strong statistical metrics such as high entropy and favorable NPCR/UACI values. Chapter~3 extended this line of analysis to the MMCBIE scheme, which employs the Hénon map, cyclic key mixing, and a two-dimensional logistic map. An efficient attack was developed that recovers the original image exactly using at most 256 oracle queries by exploiting insufficient inter-pixel and inter-channel diffusion. Both chapters together reinforce a fundamental cryptographic principle: statistical randomness does not imply security against structured active attacks.

Chapter~4 introduced PoolLiteSIS, a lightweight secret image sharing scheme that achieves a 75\% reduction in shadow image storage through pooling-based dimensional reduction. Lossless reconstruction was verified with PSNR~$= \infty$ and SSIM~$= 1.0$, and a residual permutation mechanism was incorporated to enhance security by obscuring structural relationships in the pooled shares.

Chapter~5 assessed the robustness of XOR-based visual secret sharing schemes against learning-based reconstruction attacks. Neural network adversaries were shown to achieve near-lossless reconstruction (SSIM~$= 0.9720$, PSNR~$= 33.87$~dB) when all shares are available, confirming that data-driven models can learn the full reconstruction mapping. However, under missing-share conditions, reconstruction degrades significantly, preserving the information-theoretic security guarantees of the XOR scheme.

Chapter~6 proposed an autoencoder-based secret image sharing scheme that encodes images into compact latent representations divided into independent shares. The method achieved PSNR values above 30~dB and SSIM values close to 1.0, with up to 98\% compression compared to the original image, while individual shares revealed negligible information. The trade-off is a minor reconstruction loss inherent to lossy latent-space compression.

Chapter~7 provided a systematic exploration of optimization techniques for secret image sharing, including binary representation, digit-based decomposition, average pooling with residual mapping, and XOR-based schemes with pooling. The $(t,n)$ Boolean-based MSIS scheme enabled deterministic and lossless recovery using any $t$ or more consecutive shares. The autoencoder-based model demonstrated the highest compression ratio while maintaining near-perfect reconstruction, successfully achieving the objective of generating smaller shadow images without sacrificing security or recoverability.

Taken together, these contributions advance the state of the art across anomaly detection and image security, while consistently highlighting the gap between empirical performance metrics and formal security guarantees.

% ─────────────────────────────────────────────────────────────────────────────

\section{Future Works}

The research presented in this thesis opens several promising directions for further investigation.

\subsection*{Log-Based Anomaly Detection}

The role-aware framework introduced in Chapter~1 may be extended in several ways. First, adaptive role distributions could be developed to handle evolving log formats automatically, reducing maintenance overhead in production environments. Second, dynamic role count selection—learning the optimal number of latent roles per dataset—could replace the current fixed configuration. Third, multi-modal fusion of logs, metrics, and execution traces with aligned role modeling could improve detection coverage. Hierarchical roles, such as sub-roles distinguishing transient from permanent states, represent another dimension of granularity worth exploring. Finally, cross-system transfer learning through pre-training role induction on multiple heterogeneous systems could improve generalization to unseen log formats.

\subsection*{Cryptanalysis and Chaos-Based Encryption}

The cryptanalytic results in Chapters~2 and~3 suggest several avenues for future work. Extending the chosen-plaintext attack framework to other families of chaos-based encryption schemes, including those employing three-dimensional chaotic maps or permutation–diffusion architectures with feedback, would broaden the applicability of the methodology. On the constructive side, the identified vulnerabilities—insufficient diffusion, channel independence, and deterministic keystream structure—point toward concrete design guidelines for building secure chaos-based cryptosystems. Developing a systematic evaluation framework that assesses both statistical and structural attack resistance would be a valuable contribution to the field.

\subsection*{Secret Image Sharing}

The SIS contributions across Chapters~4 through~7 suggest several unified and scheme-specific directions. For PoolLiteSIS, investigating adaptive pooling strategies that balance compression ratios against reconstruction quality for high-resolution and medical imagery is a natural next step. The vulnerability assessment in Chapter~5 motivates the design of XOR-based schemes with provable resistance to learning-based adversaries, potentially through adversarial training or share obfuscation techniques.

The autoencoder-based scheme of Chapter~6 could benefit from the integration of perceptual loss functions and generative adversarial training to reduce reconstruction artifacts. Extending the framework to video secret sharing, where temporal redundancy can be exploited for additional compression, is a particularly promising direction. For the polynomial and Boolean MSIS schemes studied in Chapter~7, achieving share size reduction while preserving the $(t,n)$ threshold and lossless recovery properties remains an open problem. Hybrid schemes that combine the compression efficiency of autoencoders with the formal security guarantees of polynomial-based methods could represent a convergence of the approaches explored throughout this thesis.

\subsection*{Unified and Cross-Cutting Directions}

Across both research areas, a recurring theme is the tension between empirical performance metrics and formal security or accuracy guarantees. Future work should continue to develop evaluation benchmarks that capture this distinction—particularly for image security, where statistical tests alone have repeatedly proven insufficient. Additionally, the deployment of the anomaly detection and image sharing schemes in resource-constrained environments such as edge devices and IoT systems remains an important practical challenge, motivating further work on computational efficiency without compromising security or detection performance.