\begin{center}
\justifying
\chapter{CONCLUSIONS AND FUTURE WORK}

This thesis investigated three complementary problems centered around representation
learning and security analysis under structured and semi-structured settings:
(i) parsing-free log anomaly detection, (ii) learning-based attacks on secure image
sharing schemes, and (iii) cryptanalysis of chaos-based medical image encryption
schemes. In all three cases, the central theme was to move beyond traditional
assumptions and evaluate whether data-driven or structural analysis methods can
reveal hidden patterns or vulnerabilities that classical approaches either ignore
or assume to be secure. :contentReference[oaicite:0]{index=0}

\section{Conclusion}

For Problem~1, this work addressed the structural limitations of existing parsing-free
log anomaly detection methods, particularly the uniform treatment of tokens and the
discarding of numeric information in NeuralLog~\cite{neurallog2021}. A role-aware
representation learning framework was proposed, which induces latent token roles —
state, entity, action, quantity, and location — directly from sequence-level anomaly
labels. The framework preserves structural and semantic distinctions through
role-aware pooling while operating directly on raw log text without requiring log
parsing.

Experimental evaluation on HDFS, BGL, Thunderbird, and Spirit datasets demonstrated
consistent improvements over traditional machine learning methods and deep learning
baselines. The proposed model achieved F1-scores in the range of 0.98--0.99 across
all datasets. Ablation studies confirmed the importance of both role modeling and
numeric token preservation. The learned roles aligned well with intuitive semantic
categories despite being induced without supervision, providing both performance
gains and interpretability.

For Problem~2, this work examined the robustness of XOR-based visual secret sharing
schemes under learning-based adversaries. Through systematic experimentation, it was
shown that neural networks can learn to approximate the reconstruction function of
XOR-based schemes when all shares are available. However, under missing-share
conditions, reconstruction quality degrades significantly, and meaningful recovery
becomes infeasible. Experiments using both four-share and three-share configurations
confirmed that the learning-based model cannot violate the fundamental
information-theoretic guarantees of the XOR scheme when insufficient shares are
provided. These results highlight that while learning-based methods can exploit data
priors, they do not inherently break cryptographic security guarantees in this
setting.

For Problem~3, this work performed a complete cryptanalysis of a chaos-based medical
image encryption scheme under the chosen-plaintext attack (CPA) model. The analysis
demonstrated that the scheme is fundamentally insecure due to structural design
flaws. Specifically, the keystream used in the diffusion stage is independent of the
plaintext, allowing it to be recovered exactly using a single chosen plaintext. In
addition, the permutation and diffusion stages are fully separable, enabling the
complete reconstruction of the permutation map using structured probe images.
Combining these results, the full encryption process can be inverted without
knowledge of the secret key.

Experimental validation confirmed that the attack requires only $n+1$ oracle queries
for an $n \times n$ image and achieves exact plaintext recovery with zero error. This
demonstrates that statistical measures such as entropy, NPCR, and UACI are not
sufficient indicators of cryptographic security. The findings highlight the critical
importance of plaintext-dependent keystream generation and tight coupling between
encryption stages in the design of secure image encryption algorithms.

\section{Future Work}

Several promising directions arise from this work.

For log anomaly detection, extending role-aware modeling to dynamic and evolving log
formats remains an important challenge. Learning adaptive role distributions and
automatically determining the optimal number of roles for different datasets could
further improve generalization. Integrating logs with other system observability data
such as metrics and traces offers a path toward more comprehensive anomaly detection.
Hierarchical role modeling may also enable finer-grained structural understanding of
complex system behaviors.

For secure image sharing, future work could explore stronger adversarial models and
more complex datasets to better understand the limits of learning-based reconstruction.
Extending analysis to higher threshold schemes and non-XOR-based constructions would
provide broader insights. Additionally, designing defenses against learning-based
attacks, such as adversarially robust share generation mechanisms, is an important
direction.

For chaos-based image encryption, future work should focus on designing schemes that
eliminate the structural weaknesses identified in this work. In particular, keystream
generation must depend on the plaintext, and permutation and diffusion stages should
be tightly coupled to prevent independent recovery. Formal cryptographic analysis,
including resistance to chosen-plaintext and chosen-ciphertext attacks, should be
incorporated alongside statistical evaluations. Exploring hybrid designs that combine
chaotic systems with established cryptographic primitives may provide stronger and
more provable security guarantees.

Overall, this work demonstrates that learning-based methods can both enhance system
analysis and expose hidden vulnerabilities in security mechanisms. It highlights the
need for integrating data-driven evaluation with rigorous structural and
cryptographic analysis in the design of modern secure systems.

\end{center}